---
title: "Generate Query from Source Table Schema & Merge w/ Manaul Query"
output: html_notebook
editor_options: 
  markdown: 
    wrap: 72
---

#### 0. Load dependencies

```{r}
library(jsonlite)
library(tools)
# Load helper functions
source("connect_analytics_helper_functions.R")
```

#### 0.5 Get schema from BigQuery

```{r}
# PROJECT="nih-nci-dceg-connect-prod-6d04"
# TABLE="Connect.participants"
# SCHEMA_JSON="participants_schema.json"
# bq show --schema --format=prettyjson $PROJECT:$TABLE > $SCHEMA_JSON

project     <- "nih-nci-dceg-connect-prod-6d04"
table       <- "Connect.module2_v2"
module      <- strsplit(table, split=".", fixed=TRUE)[[1]][2]
schema_json <- paste0(module, "-schema.json")
schema_csv  <- paste0(module, "-schema.csv")
set_default_gcp_project(project)
get_gcp_table_schema(project, table, schema_json, schema_csv) 
```

#### 1. Generate \*-variables.csv and \*-lists.json file from source table schema.

Here, we'll generate a new query from a BigQuery table schema, given as
an excel sheet which can be exported from BigQuery. Our queries require
a query generator, a variables.csv file and a lists.json file. This
function generates filters the variables in the schema based on whether
they should be in the variables file or the list file. All variables
destined for the list file are assigned a list of CIDs that are present
in the data set for that variable, across all rows.

```{r}
project     <- "nih-nci-dceg-connect-dev"
# project   <- "nih-nci-dceg-connect-stg-5519"
# project   <- "nih-nci-dceg-connect-prod-6d04"
table       <- "Connect.module2_v2"
schema_json <- "module2_v2-schema.json"
Sys.setenv(OUT_JSON = out_json)
Sys.setenv(OUT_CSV = out_csv)

schema      <- "module2_v2-schema.xlsx"
sheet       <- NULL
out_json    <- "module2_v2-lists-from-schema.json" # destination file
out_csv     <- "module2_v2-variables-from-schema.csv"

# Set output files to environment variables for easy cleanup later
Sys.setenv(OUT_JSON = out_json)
Sys.setenv(OUT_CSV = out_csv)

# Generate the query from the schema in the form of a set of 
# variables.csv and lists.json files.
out <- filter_vars_from_schema(project, table, schema, out_csv, out_json, 
                               sheet = sheet,       
                               output_file_type = "json")
out$variables[0:10]
out$array_json[0:10]
```

#### 2. Combine json and csv files prior to generating query

Now that we have a new query in the form of variables.csv and lists.json
files, we need to combine the new json file with the old json file.

```{r}
A <- "../module2_v2-lists.json"
B <- "module2_v2-lists-from-schema.json"
C <- "module2_v2-lists-union.json"

Sys.setenv(UNION_JSON = C) # For easy cleanup later

union_json_str <- get_union_of_json_arrays(A, B, C)
union_json_str
```

We also need to combine the variables.csv file that was generated from
the schema with the old, manually generated variables.csv file. This
code will exclude any duplicates.

```{r}
# Get union of variable lists
A_csv <- "../module2_v2-variables.csv"
B_csv <- "module2_v2-variables-from-schema.csv"
C_csv <- "module2_v2-variables-union.csv"

Sys.setenv(UNION_CSV = C_csv) # For easy cleanup later

union_var_chr_array <- get_union_of_variable_csv(A_csv, B_csv, C_csv)
union_var_chr_array
```

#### 3. Deal with exceptions

```{r}
json_file         <- C
variables_file    <- C_csv
# Get a list of variables with arrays containing "exception CID"
exceptions        <- c(178420302, 958239616, 353358909, 104430631)
vars_w_exceptions <- get_vars_with_specified_vals(json_file, 
                                                  exceptions, 
                                                  max_array_len = 2) 
vars_w_exceptions
```

Remove list of variables with exceptions from the data set

```{r}
output_filename  <- "module2_v2-lists-union-mod.json"
Sys.setenv(NEW_JSON = output_filename) # For easy cleanup later
updated_variables <- remove_vars_from_json(json_file, 
                                           vars_w_exceptions,
                                           output_filename = output_filename) 
updated_variables
```

Add list of variables with exceptions to \*-variables\*.csv

```{r}
out_file  <- "module2_v2-variables-union-new.csv"
Sys.setenv(NEW_CSV = out_file) # For easy cleanup 
add_novel_vars_to_csv(vars_w_exceptions, variables_file, out_file = out_file)
```

Get list of variables from csv file that are of typ RECORD or mode
REPEATED in the schema

```{r}
#TODO This chunk needs to be tested.
csv_file      <- out_file
recs_and_reps <- get_record_and_repeated_vars(project, table, schema, csv_file)
recs_and_reps
```

Remove these variables from the csv file

```{r}
#TODO This chunk needs to be tested.
in_file        <- C_csv
vars_to_remove <- recs_and_reps
remove_vars_from_csv(in_file, vars_to_remove, out_file = NULL)
```

#### 4. Move modified lists.json and variables.csv files to parent directory

```{bash}
# Move up-to-date files to parent directory
echo $NEW_CSV  
echo $NEW_JSON
PAR_DIR_CSV_NAME="module2_v2-variables.csv"
PAR_DIR_JSON_NAME="module2_v2-lists.json"

cp $NEW_CSV ../$PAR_DIR_CSV_NAME
cp $NEW_JSON ../$PAR_DIR_JSON_NAME
```

#### 5. Clean up intermediate files from working directory

```{bash}
# Delete intermediate files from directory
echo "The files that will be removed are:"
echo $OUT_JSON
echo $OUT_CSV
echo $UNION_JSON
echo $UNION_CSV
echo $NEW_CSV  
echo $NEW_JSON

rm $OUT_JSON
rm $OUT_CSV
rm $UNION_JSON
rm $UNION_CSV
rm $NEW_CSV  
rm $NEW_JSON
```

#### 6. Generate query using Warren-style Query Generator

```{bash}
# Replace name of query generator as needed
node ../module2_v2-query_generator.js
```
