---
title: "Generate Query from Source Table Schema"
output: html_notebook
editor_options: 
  markdown: 
    wrap: 72
---

Here, we'll generate a new query from a BigQuery table schema, given as
an excel sheet which can be exported from BigQuery. Our queries require
a query generator, a variables.csv file and a lists.json file. This
function generates filters the variables in the schema based on whether
they should be in the variables file or the list file. All variables
destined for the list file are assigned a list of CIDs that are present
in the data set for that variable, across all rows.

```{r}
library(jsonlite)
library(tools)
source("connect_analytics_helper_functions.R")

# Generate the query from the schema in the form of a set of variables.csv and
# lists.json files.
project  <- "nih-nci-dceg-connect-prod-6d04"
table    <- "Connect.participants"
schema   <- "/Users/petersjm/Documents/dataFlattening/autoFlattening/module_1_schemas.xlsx"
sheet    <- "M1V1"
out_json <- "M1V1-lists-auto.js" # destination file
out_csv  <- "M1V1-variables_auto.csv"

filter_vars_from_schema(project, table, schema, out_json, out_csv, sheet=NULL)

# Import the data from the files that were generated 
variables <- read.csv(file = out_csv, header = FALSE)
head(variables)
lists <- readLines(out_json, n =10)
head(lists)
```

Now that we have a new query in the form of variables.csv and lists.json
files, we need to combine the new json file with the old json file,

```{r}
A <- "M1V1-lists-auto.json"
B <- "M1-lists.json"
C <- "M1-lists-union.json"

source("connect_analytics_helper_functions.R")
union_json_str <- get_union_of_json_arrays(A, B, C)
union_json_str
```

We also need to combine the variables.csv file that was generated from
the schema with the old, manually generated variables.csv file. This
code will exculde any duplicates.

```{r}
# Get union of variable lists
A_csv <- "M1V1-variables-auto.csv"
B_csv <- "M1-variables.csv"
C_csv <- "M1V1-variables-union.csv"
union_var_chr_array <- get_union_of_variable_csv(A_csv, B_csv, C_csv)
union_var_chr_array
```

Deal with exceptions

```{r}
json_file         <- "M1-lists-union.json"
variables_file    <- "M1V1-variables-auto.csv"
# Get a list of variables with arrays containing "exception CID"
exceptions        <- c(178420302, 958239616, 353358909, 104430631)
vars_w_exceptions <- get_vars_with_specified_vals(json_file, 
                                                  exceptions, 
                                                  max_array_len = 2) 
vars_w_exceptions
```

Remove list of variables with exceptions from the data set

```{r}

output_filename  <- "M1V1-lists-union-mod.json"
updated_variables <- remove_vars_from_json(json_file, 
                                           vars_w_exceptions,
                                           write_to_file = TRUE, 
                                           output_filename = output_filename) 
updated_variables
```

Add list of variables with exceptions to \*-variables.csv

```{r}

out_file <- "M1V1-variables-auto-new.csv"
add_novel_vars_to_csv(vars_w_exceptions, variables_file, out_file = out_file)
```
